<!-- question-type: inclass -->
## Exercise 5: Developing Another Time-Based Model

The lag-1 model assumes that the **most recent game** is the most useful past game for predicting attendance at the next game. But this is only one way to use the timing of past games.

Your task is to develop a **different time-based prediction rule** using the training data.

### (a) Propose a prediction rule

Propose a prediction rule that uses attendance at previous games to predict attendance at the current game.

Your rule should use only information that would have been available **before the game you are predicting**.

### (b) Defend your choice

Explain why you think your proposed rule could produce sensible predictions of White Sox attendance. What information from recent attendance does your rule capture, and why might that be useful?

### (c) Create your predictions

Implement your prediction rule using `chw_train`.

Store your predictions in a new column. Give the column a short, descriptive name that makes it clear which model it belongs to.

Remember to:

- calculate predictions separately within each season;
- arrange games in chronological order before using information from previous games; and
- `ungroup()` once you have finished any grouped calculations.

```{r}
#| eval: false

YOUR_CODE_HERE
```

### (d) Evaluate your model

Use `attendance_metrics()` to evaluate your new model on the training data.

```{r}
#| eval: false

YOUR_CODE_HERE
```

Compare these results with the training performance of the lag-1 model.

Does your new model improve on the lag-1 model for any of the three metrics? Does it perform worse on any?


<!-- BEGIN PROFILE:r-teaching-guide -->
::: {.content-visible when-profile="r-teaching-guide"}

::: {.teaching-block}

::: {.teaching-block-header}
Teaching Note
:::

::: {.teaching-block-body}

🎯 **Learning Objective**

Students should:

- recognise that there are many reasonable ways to use recent attendance to construct a time-based prediction;
- make and defend an analytical modelling choice before writing code;
- ensure that a prediction uses only information available before the game being predicted; and
- compare the training performance of their proposed model with the lag-1 benchmark.

✅ **Core Concepts to Highlight**

- There is deliberately no single correct model in this exercise. Students should make a reasonable proposal and explain the logic behind it.
- A sensible time-based model should use only past attendance. Future attendance cannot be used to predict an earlier game.
- Encourage students to think about the trade-off between using the most recent information and smoothing over several recent games.
- Different groups may produce different models. This is useful: the exercise is about analytical judgement, not reproducing one prescribed answer.
- The illustrative solution uses the mean of the previous three games. This is only one reasonable approach and should not be presented as the model students were expected to discover.
- Students should continue to calculate lagged quantities within season and `ungroup()` afterwards.
- Comparisons in part (d) are still based on training performance. Do not let a lower training error become a claim that the new model will necessarily predict unseen games better.

💬 **Discussion Prompts**

Before students start coding, ask a few groups to share their proposed rules and reasoning. Useful questions include:

- Why might your rule be better than using only the previous game?
- What information does your rule give more or less weight to?
- How many previous games does your model need before it can make its first prediction?
- Would your rule respond quickly or slowly if attendance suddenly changed?

If students immediately converge on a three-game average, encourage alternatives rather than telling the whole room to use the same rule.

🛟 **If Students Are Stuck**

Do not immediately give students the three-game trailing-average solution. First, remind them that the lag-1 model uses only the previous game's attendance and ask: **"What else could we do with attendance from previous games?"**

If they are still stuck, offer some possibilities to choose from:

- take the **average of the previous two games**;
- take the **average of several previous games**;
- give **more weight to recent games** and less weight to older games;
- use a **longer window of recent games** to produce a smoother prediction.

A model based simply on attendance two games ago is also possible, but it is less interesting here: it mainly changes which single past game is used rather than combining information from recent games.

Ask the group to choose an approach and explain why it might be useful **before** they start coding. The goal is not to steer everyone toward the same model, but to help students turn an intuitive idea about recent attendance into a prediction rule they can defend.

If students struggle to implement their chosen idea, help them translate the rule into the required `lag()` operations rather than replacing their choice with the illustrative three-game-average solution.

📌 **Common Misunderstandings**

- Students may calculate a moving average that includes the current game's attendance. That would use the outcome they are trying to predict.
- Students may forget to arrange games chronologically before creating lagged variables.
- Students may allow the end of one season to feed into predictions for the next season.
- Students may propose a rule but give no substantive reason why it could help prediction.
- Different models will have different numbers of missing predictions at the beginning of each season. For example, a three-game trailing average cannot make a prediction until three previous games are available.
- Students may interpret improved training performance as proof that their model is better for future games. Reinforce that test performance is still required.

:::

:::

:::
<!-- END PROFILE:r-teaching-guide -->

<!-- BEGIN PROFILE:r-solutions -->
::: {.content-visible when-profile="r-solutions" when-profile="r-teaching-guide"}

::: {.solution-block}

::: {.solution-block-header}
Solution
:::

::: {.solution-block-body}

There are many reasonable answers to this exercise. The solution below illustrates one possible approach: predicting attendance using the **average attendance at the previous three games**.

#### (a)

One possible prediction rule is to use the mean attendance from the previous three White Sox home games in the same season.

#### (b)

Using three previous games may produce a more stable prediction than using only the immediately preceding game. Attendance at a single game could be unusually high or low for reasons that do not persist to the next game. Averaging several recent games smooths some of this game-to-game variation while still using relatively recent information.

There is a trade-off. Using several games means the prediction may respond more slowly if attendance patterns change quickly, and the model cannot make a prediction until enough previous games are available.

#### (c)

```{r}
chw_train_pred_trailing3 <-
  chw_train |>
  group_by(year) |>
  arrange(gm, .by_group = TRUE) |>
  mutate(
    pred_trailing3 = (
      lag(attendance, 1) +
      lag(attendance, 2) +
      lag(attendance, 3)
    ) / 3
  ) |>
  ungroup()
```

This calculation uses only attendance from the three games that occurred before the game being predicted. The first three games of each season therefore have missing predictions.

#### (d)

```{r}
chw_train_pred_trailing3 |>
  attendance_metrics(
    truth = attendance,
    estimate = pred_trailing3
  )
```

The exact comparison with the lag-1 model depends on the resulting training metrics. Students should identify which metrics improve or worsen relative to their lag-1 results, while remembering that this is still a comparison on the training data.

:::

:::

:::
<!-- END PROFILE:r-solutions -->
