<!-- question-type: prepare -->
## Exercise 2: Predicting Attendance Using the Previous Game

One simple way to predict attendance at a future game is to use attendance at a recent game. We will start with a **lag-1 model**: our prediction for each game is the attendance at the White Sox's previous home game.

### (a) Create lag-1 predictions

Starting with `chw_train`, create a lag-1 prediction of attendance for each game.

Your code should:

- group games by season;
- arrange games in chronological order;
- create `pred_lag1`, containing attendance at the previous game;
- ungroup the data; and
- keep only `gm`, `year`, `attend`, and `pred_lag1`.

Store the result in `chw_train_pred`.

```{r}
#| eval: FALSE
chw_train_pred <-
  YOUR_CODE_HERE |>
  select(gm, year, attend, pred_lag1)
```

### (b) Why do we group by season?

Suppose we instead arranged **all** games from 2022 to 2024 chronologically and then used `lag(attend)` without first grouping by season.

What would the lag-1 prediction be for the **first game of the 2024 season**? Why would that not be a sensible prediction for our model?

### (c) What happens to the first game?

Inspect the first few games from each season in `chw_train_pred`.

Why is `pred_lag1` missing for the first game of each season? Is this what you would expect given how we have defined the model?

### (d) Evaluate the lag-1 model

We will use **MAE**, **RMSE**, and **MAPE** to evaluate attendance predictions. Create a metric set containing these three measures and store it as `attendance_metrics`.

Then use `attendance_metrics()` to evaluate the lag-1 model on the training data.

```{r}
#| eval: FALSE

attendance_metrics <- YOUR_CODE_HERE

chw_train_pred |>
  attendance_metrics(
    truth = YOUR_CODE_HERE,
    estimate = YOUR_CODE_HERE
  )
```



<!-- BEGIN PROFILE:r-teaching-guide -->
::: {.content-visible when-profile="r-teaching-guide"}

::: {.teaching-block}

::: {.teaching-block-header}
Teaching Note
:::

::: {.teaching-block-body}

🎯 **Learning Objective**

Students should:

- construct a simple time-based prediction using `lag()`;
- understand why observations must be ordered chronologically before creating a lag;
- understand why the lag should be calculated within season; and
- evaluate training predictions using MAE, RMSE, and MAPE.

✅ **Core Concepts to Highlight**

- A lag-1 model uses the attendance at the previous home game in the same season as the prediction for the current game.
- `arrange()` determines what "previous" means before `lag()` is applied.
- `group_by(year)` prevents the final game of one season from becoming the prediction for the first game of the next season.
- `ungroup()` is important once the grouped calculation is complete so that the grouping does not unexpectedly affect later operations.
- The first game of each season has no lag-1 prediction because there is no previous game in that season.
- These metrics describe performance on the training data. They do not yet tell us how well the model predicts unseen games.

📌 **Common Misunderstandings**

- Students may use `lag()` without first arranging the games chronologically.
- Students may forget to group by `year`, creating a prediction for the first game of a season from the previous season's final game.
- Students may forget to `ungroup()`.
- Students may reverse `truth` and `estimate` when calling `attendance_metrics()`.
- Students may interpret the training metrics as evidence that this is a good model for future games. Save that discussion for the in-class comparison of training and test performance.

:::

:::

:::
<!-- END PROFILE:r-teaching-guide -->

<!-- BEGIN PROFILE:r-solutions -->
::: {.content-visible when-profile="r-solutions" when-profile="r-teaching-guide"}

::: {.solution-block}

::: {.solution-block-header}
Solution
:::

::: {.solution-block-body}

#### (a)

```{r}
chw_train_pred <-
  chw_train |>
  group_by(year) |>
  arrange(gm, .by_group = TRUE) |>
  mutate(
    pred_lag1 = lag(attendance)
  ) |>
  ungroup() |>
  select(gm, year, attendance, pred_lag1)
```

#### (b)

Without grouping by season, the lag-1 prediction for the first game of the 2024 season would be attendance at the final game of the 2023 season. This is not the prediction rule we want: the model is intended to use the previous home game **within the same season**.

#### (c)

`pred_lag1` is missing for the first game of each season because there is no previous game within that season. This is exactly what we should expect from the way the lag-1 model is defined.

#### (d)

```{r}
attendance_metrics <-
  metric_set(mae, rmse, mape)

chw_train_pred |>
  attendance_metrics(
    truth = attendance,
    estimate = pred_lag1
  )
```

:::

:::

:::
<!-- END PROFILE:r-solutions -->
