<!-- question-type: inclass -->
### Exercise 7: How Well Do Our Models Predict New Games?

We developed our prediction rules using the training data. Now we can see how well they predict **new games**.

#### (a)

Apply the prediction rules you developed in the previous exercises to the 2025 test data. Do not use 2025 attendance to construct or update your prediction rules.

Store the predictions together with actual attendance in `chw_test_preds`.

```{r}
#| eval: false

chw_test_preds <-
  chw_test |> 
  YOUR_CODE_HERE
```

#### (b) Compare the same games

Create `chw_test_complete` by keeping only games for which all of the models you want to compare have predictions. Why does evaluating the models on the **same games** make the comparison fairer?


```{r}
#| eval: false

chw_test_complete <-
  chw_test_preds |>
  filter(
    YOUR_CODE_HERE
  )
```


#### (c) Reshape the predictions

Each model is currently stored in a different prediction column. Explain why this **wide format would lead us to repeat very similar code** when evaluating several models.

#### (d)

Rather than keeping the predictions in wide format, reshape them into long format.

Use `pivot_longer()` to create a `model` column and a `.pred` column.

```{r}
#| eval: false

chw_test_long <-
  chw_test_complete |>
  pivot_longer(
    cols = YOUR_CODE_HERE,
    names_to = YOUR_CODE_HERE,
    values_to = YOUR_CODE_HERE
  )
```

#### (e) Evaluate the models

Calculate MAE, RMSE, and MAPE for each model.

```{r}
#| eval: false

test_metrics <-
  chw_test_long |>
  YOUR_CODE_HERE |>
  attendance_metrics(
    truth = YOUR_CODE_HERE,
    estimate = YOUR_CODE_HERE
  )

test_metrics
```

#### (f) Make the results easier to compare

Use `pivot_wider()` to reshape the data once more to create one row per model, with separate columns for MAE, RMSE, and MAPE.

```{r}
#| eval: false

test_metrics_wide <-
  test_metrics |>
  select(model, .metric, .estimate) |>
  pivot_wider(
    names_from = YOUR_CODE_HERE,
    values_from = YOUR_CODE_HERE
  )

test_metrics_wide
```

#### (g) 

Which is the best prediction model? Explain your reasoning.

<!-- BEGIN PROFILE:r-teaching-guide -->
::: {.content-visible when-profile="r-teaching-guide"}

::: {.teaching-block}

::: {.teaching-block-header}
Teaching Note
:::

::: {.teaching-block-body}

🎯 **Learning Objective**

Students should:

- apply prediction rules developed using the training data to unseen test data;
- understand why models should be compared on the same observations;
- use long-format data to evaluate several prediction models efficiently;
- use wide-format data to compare model performance; and
- select a prediction model using test-set performance.

✅ **Core Concepts to Highlight**

- The 2025 data are used to evaluate prediction rules developed using the training data.
- Comparing models on the same games ensures differences in metrics are not driven by different evaluation samples.
- `pivot_longer()` turns the prediction model from separate columns into a `model` variable, allowing the same metric calculation to be applied across models.
- `pivot_wider()` makes the resulting metrics easier to compare across models.
- Long and wide formats are useful for different analytical tasks rather than one format being inherently better.
- Test performance is the relevant evidence when choosing among models intended to predict new games.
- In the provided solution, `pred_opp_lag1` performs best on all three test metrics:
  - MAE = 4,393;
  - RMSE = 6,386; and
  - MAPE = 25.1%.

📌 **Common Misunderstandings**

- Students may use 2025 attendance to re-estimate a prediction rule rather than only to evaluate predictions.
- Students may compare models using different sets of games when some predictions are missing.
- Students may repeat the metric calculation separately for each prediction column rather than using the long format and `group_by(model)`.
- Students may think long-format data are always preferable to wide-format data, rather than recognising that the useful shape depends on the task.
- Students may choose a model based on training performance rather than its performance on unseen test data.
- Students may select a model using only one metric without noticing that, in this example, all three metrics identify the same best-performing model.


:::

:::

:::
<!-- END PROFILE:r-teaching-guide -->

<!-- BEGIN PROFILE:r-solutions -->
::: {.content-visible when-profile="r-solutions" when-profile="r-teaching-guide"}

::: {.solution-block}

::: {.solution-block-header}
Solution
:::

::: {.solution-block-body}

There is no single solution to part (a) because students may have developed different prediction rules.

The solution below illustrates one path using:

- lag-1;
- a trailing-three average;
- average attendance by opponent; and
- attendance at the previous game against the same opponent.

#### (a)

First create the time-based predictions for the 2025 games:

```{r}
chw_test_preds <-
  chw_test |>
  arrange(gm) |>
  mutate(
    pred_lag1 = lag(attendance),
    pred_trailing3 = (
      lag(attendance, 1) +
      lag(attendance, 2) +
      lag(attendance, 3)
    ) / 3
  )
```

For the opponent-average model, calculate the prediction using only the training data:

```{r}
attendance_by_opp <-
  chw_train |>
  group_by(opp) |>
  summarise(
    pred_opp = mean(attendance, na.rm = TRUE),
    .groups = "drop"
  )
```

For the previous-same-opponent model, calculate the lag while the historical games are still available:

```{r}
chw_opp_lag1 <-
  chw_attend |>
  arrange(year, gm) |>
  group_by(opp) |>
  mutate(
    pred_opp_lag1 = lag(attendance)
  ) |>
  ungroup() |>
  filter(year == 2025) |>
  select(gm, pred_opp_lag1)
```

Then add the opponent-based predictions:

```{r}
chw_test_preds <-
  chw_test_preds |>
  left_join(
    attendance_by_opp,
    by = "opp"
  ) |>
  left_join(
    chw_opp_lag1,
    by = "gm"
  )
```

Students using different rules should instead apply the rules they developed in the previous exercises.

#### (b)

For the illustrative models:

```{r}
chw_test_complete <-
  chw_test_preds |>
  filter(
    !is.na(pred_lag1),
    !is.na(pred_trailing3),
    !is.na(pred_opp),
    !is.na(pred_opp_lag1)
  )

nrow(chw_test_complete)
```

Evaluating the models on the same games means differences in the performance metrics are not driven by the models being evaluated on different sets of observations.

#### (c)

In the wide data, each model is stored in a different prediction column. Calculating the same metrics for each model would therefore require us to repeat very similar code while changing only the prediction column.

#### (d)

```{r}
chw_test_long <-
  chw_test_complete |>
  pivot_longer(
    cols = starts_with("pred_"),
    names_to = "model",
    values_to = ".pred"
  )
```

The model names that previously appeared across several columns are now values of the `model` variable. All predictions are stored in the `.pred` variable.

#### (e)

```{r}
test_metrics <-
  chw_test_long |>
  group_by(model) |>
  attendance_metrics(
    truth = attendance,
    estimate = .pred
  )

test_metrics
```

Because model identity is now a variable, `group_by(model)` allows the same call to `attendance_metrics()` to be applied to every model.

#### (f)

```{r}
test_metrics_wide <-
  test_metrics |>
  select(model, .metric, .estimate) |>
  pivot_wider(
    names_from = .metric,
    values_from = .estimate
  )

test_metrics_wide
```

#### (h)

`pred_opp_lag1` is the best-performing prediction model. It has the lowest MAE, RMSE, and MAPE on the test data. 


:::

:::

:::
<!-- END PROFILE:r-solutions -->
