<!-- question-type: prepare -->
## Exercise 3: Predicting Attendance Using the Opponent

Our lag-1 model uses **when a game occurs** to decide which past game should inform our prediction. Another approach is to look for past games that are similar in some other way.

One potentially useful characteristic is the **opponent**. Some visiting teams may consistently attract larger crowds than others.

For each opponent, we will use their average attendance in **2023** as our prediction for games against that opponent.

### (a) Calculate average attendance by opponent

Using `chw_train`, keep only games from 2023. Then:

- group the data by `opp`;
- calculate mean `attendance` for each opponent; and
- store the mean in a variable called `pred_opp`.

Store the resulting table as `attendance_by_opp`.

```{r}
#| eval: false
attendance_by_opp <-
  YOUR_CODE_HERE
```

### (b) Create predictions for the training data

We now have one predicted attendance value for each opponent.

Use a `left_join()` to add `pred_opp` to `chw_train`. Store the result as `chw_train_pred_opp`.

```{r}
#| eval: false
chw_train_pred_opp <-
  YOUR_CODE_HERE |>
  select(gm, year, opp, attendance, pred_opp)
```

Why do we use `left_join()` with `chw_train` on the left-hand side?

### (c) Evaluate the opponent model

Use `attendance_metrics()` to evaluate the opponent model on the training data.

```{r}
#| eval: false
chw_train_pred_opp |>
  attendance_metrics(
    truth = YOUR_CODE_HERE,
    estimate = YOUR_CODE_HERE
  )
```


<!-- BEGIN PROFILE:r-teaching-guide -->
::: {.content-visible when-profile="r-teaching-guide"}

::: {.teaching-block}

::: {.teaching-block-header}
Teaching Note
:::

::: {.teaching-block-body}

🎯 **Learning Objective**

Students should:

- understand a group-based prediction rule as an alternative to a time-based rule;
- create a prediction by calculating historical mean attendance within opponent groups;
- apply a prediction table to other observations using `left_join()`; and
- evaluate the resulting predictions on the training data.

✅ **Core Concepts to Highlight**

- The opponent model defines "similar games" as games against the same opponent.
- The prediction rule is learned from 2023 games: each opponent receives its mean 2023 attendance.
- `summarise()` creates a lookup table with one prediction for each opponent.
- `left_join()` applies that lookup table to the training games by matching on `opp`.
- Keeping `chw_train` on the left of the join preserves the games we want to predict.
- As with the lag-1 model, these metrics describe performance on the training data. We have not yet evaluated performance on unseen games.

📌 **Common Misunderstandings**

- Students may calculate opponent means using all years rather than filtering to 2023 first.
- Students may forget to group by `opp` before calculating mean attendance.
- Students may join in the wrong direction or omit the join key.
- Some opponents may not have a 2023 prediction available. If this occurs, use it as an opportunity to discuss that a group-based prediction can only be applied when the relevant historical group exists.
- Students may compare the training metrics and immediately conclude that one model is better for predicting future games. Reinforce that this comparison is not yet based on unseen data.

:::

:::

:::
<!-- END PROFILE:r-teaching-guide -->

<!-- BEGIN PROFILE:r-solutions -->
::: {.content-visible when-profile="r-solutions" when-profile="r-teaching-guide"}

::: {.solution-block}

::: {.solution-block-header}
Solution
:::

::: {.solution-block-body}

#### (a)

```{r}
attendance_by_opp <-
  chw_train |>
  filter(year == 2023) |>
  group_by(opp) |>
  summarise(
    pred_opp = mean(attendance)
  )
```

#### (b)

```{r}
chw_train_pred_opp <-
  chw_train |>
  left_join(attendance_by_opp, by = "opp") |>
  select(gm, year, opp, attendance, pred_opp)
```

We use `left_join()` with `chw_train` on the left-hand side because we want to keep the White Sox training games and add the corresponding opponent-based prediction to each game.

#### (c)

```{r}
chw_train_pred_opp |>
  attendance_metrics(
    truth = attendance,
    estimate = pred_opp
  )
```

:::

:::

:::
<!-- END PROFILE:r-solutions -->
